---
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  echo = TRUE,
  comment = "#>",
  out.width = "100%")

## read in packages

library(knitr)
library(tidyverse)


## baseline model
load("baseline_model.Rdata")


#knit(input="riskscore_paper.rmd", output = "readme.md") 
```

# Diabetes riskscore (additional files)

**For the visualisation of the interaction, please see: [https://kristyrobledo.github.io/diabetes_riskscore/](https://kristyrobledo.github.io/diabetes_riskscore/)**

## Model explanation

The `baselinemodel_age.Rdata` object is an R model object for the final model, including the following baseline risk factors: 2-hour OGTT glucose, HbA1c,  age and treatment with testosterone

**Outcome**

- `ogtt_gt11` = factor. 0= “no diabetes at 2 years”, 1=“diabetes at 2
  years”, as determined by 2-hour OGTT glucose \>=11.1mmol/l

**Covariates**

- `cxhba1c` = numeric. Hba1c level at baseline in %.
- `cxglu2h` = numeric. 2-hour OGTT glucose at baseline in mmol/L ,
- `age` = numeric. age at baseline in years (may have decimal places)
- `t_x` = numeric. Interaction term with treatment and hba1c, see
  example for how to calculate. Briefly: if planned to not have testosterone treatment (ie. placebo), always 0. If planning to treat with Testosterone, and hba1c values is \<5.6%, its 0. If Hba1c is \>5.6%,
  its the amount that the hba1c value is \>5.6%.

## Example of use

To use the risk score, let create some fake patients to use this on. The
most complex part is to create the interaction variable, but this is straightforward
with the code below:

```{r}
library(tidyverse)

# setup hba1c and glucose values for 10 patients
cxhba1c <- seq(min(4.1), max(6.4), length.out = 10)
cxglu2h <- seq(min(7.7), max(11), length.out = 10)

# create a dataset with 10 placebo (treat=0) and 10 testosterone (treat=1) patients aged 59 years
as.data.frame(
  rbind(
    cbind(age=59,  cxhba1c=cxhba1c, cxglu2h = cxglu2h, treat=rep(0,10)), 
    cbind(age=59, cxhba1c=cxhba1c, cxglu2h = cxglu2h, treat=rep(1,10))
  )
) ->pred_df

#glimpse(pred_df)

##create the interaction variable
pred_df %>%
  mutate(
    t_x = case_when(treat==0 ~0,
                treat==1 & cxhba1c<5.6  ~0,
                treat==1 & cxhba1c>=5.6~ cxhba1c-5.6)) ->pred_df

#glimpse(pred_df)
head(pred_df)
tail(pred_df)
```

Now we have our prediction dataset, we can calculate the probability of diabetes at 2 years for these patients (`fit`, and the standard error called `fit.se`):

```{r}
# load the baseline model, saved in the same spot as this file
load("baselinemodel_age.Rdata")

prob_df<-cbind(pred_df, predict(baselinemodel_age,
                                type="response",
                                se.fit=TRUE,
                                newdata = pred_df))

head(prob_df)
tail(prob_df)
```


