---
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  echo = TRUE,
  comment = "#>",
  out.width = "100%")

## read in packages

library(knitr)
library(tidyverse)


## baseline model
load("baseline_model.Rdata")


#knit(input="riskscore_paper.rmd", output = "readme.md") 
```

# Diabetes riskscore (additional files)


## Model objects (from R)

The R model objects for the three models are as follows:

- `baseline_model.Rdata` = baseline risk factors: 2-hour OGTT glucose, HbA1c and treatment with testosterone
- `baseline_weight_model.Rdata` = baseline risk factors plus baseline weight and changes in weight at one year
- `baseline_waist_model.Rdata` = baseline risk factors plus baseline waist circumference and changes in waist circumference at one year

## Names and units

### baseline_model

**Outcome**

- `ogtt_gt11` = factor. 0= "no diabetes at 2 years", 1="diabetes at 2 years", as determined by 2-hour OGTT glucose >=11.1mmol/l 

**Covariates**

- `cxhba1c` = numeric. Hba1c level at baseline in %. 
- `cxglu2h` = numeric. 2-hour OGTT glucose at baseline in mmol/L , 
- `t_x` = numeric. Interaction term with treatment and hba1c, see example for how to calculate. Briefly: if placebo, always 0. If Testosterone, and hba1c values is <5.6%, its 0. If Hba1c is >5.6%, its the amount that the hba1c value is >5.6%. 

### baseline_weight_model

Model looks like so: `outcome ~ logit + base_wt + change_wt`

**Outcome**

- `outcome` = factor. 0= "no diabetes at 2 years", 1="diabetes at 2 years", as determined by 2-hour OGTT glucose >=11.1mmol/l 

**Covariates**

- `logit` = numeric. The linear predictor from the `baseline_model`
- `base_wt` = numeric. Weight at baseline in kilograms, 
- `change_wt` = numeric. Change in weight at one year from baseline, in kilograms (ie. weight at one year minus weight at baseline)

### baseline_waist_model

Model looks like so: `outcome ~ logit + base_wc + change_wc`

**Outcome**

- `outcome` = factor. 0= "no diabetes at 2 years", 1="diabetes at 2 years", as determined by 2-hour OGTT glucose >=11.1mmol/l 

**Covariates**

- `logit` = numeric. The linear predictor from the `baseline_model`
- `base_wc` = numeric. Waist circumference at baseline in centimetres, 
- `change_wc` = numeric. Change in Waist circumference at one year from baseline, in centimetres (ie. waist at one year minus waist at baseline)

## Example of use

To use the risk score, let create some fake patients to use this on. The hardest part is to create the interaction variable - straightforward with the code below:

```{r}

# setup hba1c and glucose values for 10 patients
cxhba1c <- seq(min(4.1), max(6.4), length.out = 10)
cxglu2h <- seq(min(7.7), max(11), length.out = 10)

# create a dataset with 10 placebo (treat=0) and 10 testosterone (treat=1) patients 
as.data.frame(
  rbind(
    cbind( cxhba1c=cxhba1c, cxglu2h = cxglu2h, treat=rep(0,10)), 
    cbind( cxhba1c=cxhba1c, cxglu2h = cxglu2h, treat=rep(1,10))
  )
) ->pred_df

glimpse(pred_df)

##create the interaction variable
pred_df %>%
  mutate(
    t_x = case_when(treat==0 ~0,
                treat==1 & cxhba1c<5.6  ~0,
                treat==1 & cxhba1c>=5.6~ cxhba1c-5.6)) ->pred_df

glimpse(pred_df)
head(pred_df)
tail(pred_df)
```

Now we have our prediction dataset, we can calculate the probability of diabetes at 2 years for these patients (`fit`, and the standard error called `fit.se`):

```{r}
# load the baseline model, saved in the same spot as this file
load("baseline_model.Rdata")

prob_df<-cbind(pred_df, predict(baseline_model,
                                type="response",
                                se.fit=TRUE,
                                newdata = pred_df))

head(prob_df)
tail(prob_df)
```

To apply to the models that include changes in body composition, we follow the same process. 
First lets create our dataset of patients:

```{r}
cxhba1c <- seq(min(4.1), max(6.4), length.out = 10)
cxglu2h <- seq(min(7.7), max(11), length.out = 10)
base_wt <- seq(min(65), max(170), length.out = 10)
change_wt <- seq(min(-50), max(16), length.out = 10)

# create a dataset with 10 placebo (treat=0) and 10 testosterone (treat=1) patients 
as.data.frame(
  rbind(
    cbind( cxhba1c=cxhba1c, cxglu2h = cxglu2h, 
           base_wt=base_wt,change_wt=change_wt, treat=rep(0,10)), 
    cbind( cxhba1c=cxhba1c, cxglu2h = cxglu2h,
           base_wt=base_wt,change_wt=change_wt, treat=rep(1,10))
  )
) ->predch_df

glimpse(predch_df)

##create the interaction variable
predch_df %>%
  mutate(
    t_x = case_when(treat==0 ~0,
                treat==1 & cxhba1c<5.6  ~0,
                treat==1 & cxhba1c>=5.6~ cxhba1c-5.6)) ->predch_df

glimpse(predch_df)
head(predch_df)
tail(predch_df)
```

Now we have our prediction dataset, we can calculate the linear predictor for the baseline model, using the same process as before. This gives `fit` which is our logit:

```{r}

predch_df_logit<-cbind(predch_df, 
                       logit = predict(baseline_model,
                                       newdata = predch_df))

```

Now we can perform the prediction for the baseline model with weight included:

```{r}
load("baseline_weight_model.Rdata")

probch_df<-cbind(predch_df_logit, 
                 predict(baseline_weight_model,
                                    type="response",
                                    se.fit=TRUE,
                                    newdata = predch_df_logit))
head(probch_df)
tail(probch_df)
```





